### 2013-09-28 manuscripts/dpp-msbayes/dpp-msbayes.tex
Finishing draft of results for dpp ms.

\subsection*{Power analyses: model-choice}

>>>>

\subsection*{Power analyses: model-choice}
There is a large amount of heterogeneity in estimating the number of divergence
events (\divTimeNum) both among the prior models, and across the different
series of data models (Table \ref{tabPowerModels}) (Figs.\
\labelcref{figPowerPsiExpDPP,figPowerPsiExpUniform,figPowerPsiExpOld,figPowerPsiExpUshaped,figPowerPsiUniformDPP,figPowerPsiUniformUniform,figPowerPsiUniformOld,figPowerPsiUniformUshaped,figPowerPsiOldDPP,figPowerPsiOldUniform,figPowerPsiOldOld,figPowerPsiOldUshaped}).
The \modelDPP model tends to infer highly clustered divergence models across
all three series of data models when the divergences are recent, but performs
the best of the four prior models  when divergences are random over a timescale
of about one to two coalescent units (Figs.\
\labelcref{figPowerPsiExpDPP,figPowerPsiUniformDPP,figPowerPsiOldDPP,}).
The \modelUniform model performs the best when divergences are random
over the most recent timescales, but tends to only infer a tight
cluster of \divTimeNum values even when divergences are random
over longer periods (Figs.\
\labelcref{figPowerPsiExpUniform,figPowerPsiUniformUniform,figPowerPsiOldUniform,}).
The \modelOld model performs the worst of the four prior models across
all three series of data models, inferring a single divergence event across
most of the 18,000 simulations (Figs.\
\labelcref{figPowerPsiExpOld,figPowerPsiUniformOld,figPowerPsiOldOld,}).
Even when the simulated data are identically distributed as the \modelOld model
except for the divergence times, the \modelOld is still strongly biased towards
a single event, whereas the \modelDPP model is the least likely to infer a
single event when applied to these data.
Using exponential priors on nuisance parameters does increase the power
of the \modelUshaped model compared \modelOld across all three series
of data models, but the U-shaped prior still prevents the model
from performing as well as the \modelDPP and \modelUniform models (Figs.\
\labelcref{figPowerPsiExpUshaped,figPowerPsiUniformUshaped,figPowerPsiOldUshaped,}).

The improved power of the new models are even more pronounced when looking at
estimates of the dispersion index of divergence times (\divTimeDispersion)
across the simulations (Figs.\
\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionExpOld,figPowerDispersionExpUshaped,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionUniformOld,figPowerDispersionUniformUshaped,figPowerDispersionOldDPP,figPowerDispersionOldUniform,figPowerDispersionOldOld,figPowerDispersionOldUshaped}).
The \modelDPP and \modelUniform models perform similarly across all three
series of data models, inferring values of \divTimeDispersion consistent with
one divergence event ($\divTimeDispersion < 0.01)$ in almost none of the
replicates across all the simulations (Figs.\
\labelcref{figPowerDispersionExpDPP,figPowerDispersionExpUniform,figPowerDispersionUniformDPP,figPowerDispersionUniformUniform,figPowerDispersionOldDPP,figPowerDispersionOldUniform}).
The \modelOld model infers values consistent with a single divergence event in
most of the replicates across all the simulations (Figs.\
\labelcref{figPowerDispersionExpOld,figPowerDispersionUniformOld,figPowerDispersionOldOld}).
Using exponential priors on nuisance parameters greatly increases the power
of the \modelUshaped model relative to \modelOld, but it still has
less power than the models with DPP or uniform priors across
divergence models (Figs.\
\labelcref{figPowerDispersionExpUshaped,figPowerDispersionUniformUshaped,figPowerDispersionOldUshaped}).

When looking at the estimated posterior probability of the single-divergence
model across the power analyses, we also see the increased power of the
new models (Figs.\
\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbExpUniform,figPowerPsiProbExpOld,figPowerPsiProbExpUshaped,figPowerPsiProbUniformDPP,figPowerPsiProbUniformUniform,figPowerPsiProbUniformOld,figPowerPsiProbUniformUshaped,figPowerPsiProbOldDPP,figPowerPsiProbOldUniform,figPowerPsiProbOldOld,figPowerPsiProbOldUshaped}).
The \modelDPP model estimates low posterior probability of
$\divTimeNum = 1$ across all of the simulations (Figs.\
\labelcref{figPowerPsiProbExpDPP,figPowerPsiProbUniformDPP,figPowerPsiProbOldDPP}).
The \modelDPP does infer strong support for one divergence as gauged by a Bayes
factor of greater than 10.
However, this is not that unexpected given the low prior probability of a
single divergence under this model.
Also, the Bayes factor calculations for the \modelDPP replicates are only
approximate; they are calculated base on the prior mean of the concentration
parameter \concentrationParam.

The \modelUniform model infers very low posterior probabilities for the
one-divergence model across all the simulations (Figs.\
\labelcref{figPowerPsiProbExpUniform,figPowerPsiProbUniformUniform,figPowerPsiProbOldUniform}).
The\modelOld model infers high posterior probabilities of a single divergence
for most replicates across all simulations (Figs.\
\labelcref{figPowerPsiProbExpOld,figPowerPsiProbUniformOld,figPowerPsiProbOldOld}).
Again, we see that the exponential priors on nuisance parameters greatly
increase power and result in lower estimates of the probability of
one divergence when comparing \modelUshaped to \modelOld (Figs.\
\labelcref{figPowerPsiProbExpUshaped,figPowerPsiProbUniformUshaped,figPowerPsiProbOldUshaped}).

Lastly, when we look at the estimated posterior probability of
\divTimeDispersion being consistent with one divergence parameter
($p(\divTimeDispersion < 0.01 | \ssSpace)$), we see the same pattern of model
behavior, with \modelDPP and \modelUniform inferring low probabilities across
all simulations, \modelOld inferring high probabilities, and \modelUshaped
inferring intermediate values (Figs.\
\labelcref{figPowerDispersionProbExpDPP,figPowerDispersionProbExpUniform,figPowerDispersionProbExpOld,figPowerDispersionProbExpUshaped,figPowerDispersionProbUniformDPP,figPowerDispersionProbUniformUniform,figPowerDispersionProbUniformOld,figPowerDispersionProbUniformUshaped,figPowerDispersionProbOldDPP,figPowerDispersionProbOldUniform,figPowerDispersionProbOldOld,figPowerDispersionProbOldUshaped}).
